---
title: "OPT-Neuro Motion Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE, error=FALSE, message = FALSE)
```

```{r}
library(plyr)
library(tidyverse)
library(jsonlite)
library(lubridate)
library(gridExtra)
library(reshape2)
library(kableExtra)
```


```{r}
#params_init=read.csv(file='/scratch/gherman/mriqc/bold_init.csv',header=TRUE, stringsAsFactors = FALSE, na.strings=c("NA", '', ' '),colClasses=c("session_id"="character"))

params=read.csv(file='/home/gabi/Documents/kimel/opt_qc/group_bold.tsv',header=TRUE, stringsAsFactors = FALSE, na.strings=c("NA", '', ' '),colClasses=c("session_id"="character"), sep="\t") %>% separate(data = ., col = bids_name, into = c("subject_id", "session_id", "task_id", "acq_id","run_id","type"), sep='_') %>% mutate(., subject_id=substr(subject_id, 5, 14),  session_id=substr(session_id,5, 7), acq_id=substr(acq_id, 5,7), task_id=substr(task_id,6,10))



#params=read.csv(file='/home/gabi/Documents/kimel/opt_qc/bold.csv',header=TRUE, stringsAsFactors = FALSE, na.strings=c("NA", '', ' '),colClasses=c("session_id"="character"))


params <- params %>% mutate(., mr_id=paste("OPT01",acq_id, substr(subject_id, 4, nchar(subject_id)),session_id,  sep="_"), acq_site=substr(acq_id, 1, 2)) 

rc <- read.csv(file='/home/gabi/Documents/kimel/opt_qc/OPTIMUMMainDatabaseF_DATA_2019-02-27_1329.csv', stringsAsFactors = FALSE, na.strings=c("NA",'', ' '), check.names = FALSE) %>% select(., record_id, redcap_event_name, redcap_data_access_group,mr_date, mr_id) %>% filter(., redcap_event_name=="baseline_arm_6"|redcap_event_name=="6_mth_fu_arm_6", !is.na(mr_id)) %>% mutate(., mr_id=toupper(substr(mr_id, 1, nchar(mr_id)-3)))

combo <- left_join(params, rc) %>% filter(., task_id == "rest")

combo$mr_date <- as.Date(combo$mr_date)

combo <- combo %>% mutate(., date_beforeaft=ifelse(mr_date<'2018-12-14', "before","after")) %>% filter(!is.na(date_beforeaft)) %>% mutate(., scanner=ifelse(acq_id=="CU1"|acq_id=="CU2", "CU1/2", acq_id))

```

This summarizes the MRIQC metrics for framewise displacement, a measure of motion, in OPT-Neuro. It is only examining resting state fMRI scans. This document does not report the quality of structural scans. 


The chart below shows how the percentage of scan timepoints above the 0.5mm threshold has changed over time. Each dot represents one run in the scanner - a participant usually has two runs per timepoint. The vertical line represents the date of the RA call to discuss strategies to mitigate in-scanner motion. The horizontal red line at 50% represents a liberal cut-off for when data are unusable (i.e., points that lie above the red line will likely be unsuitable for analysis). 

```{r}
xstart=as.Date("2017-11-18")
xend = as.Date("2018-12-14")

xstart2=xend
#xend2=as.Date(max(combo$mr_date))
xend2=as.Date("2019-02-20")

p <-  ggplot(data=combo, mapping = (aes(mr_date,fd_perc, col=acq_id)))+  
  annotate("rect", xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.5)  +
  annotate("rect", xmin = xstart2, xmax = xend2, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.8)  + 
  geom_point() + 
  geom_vline(xintercept=as.Date('2018-12-14'), linetype=3) + 
 geom_hline(yintercept = 50, linetype=1, colour="red") + 
  facet_wrap(~scanner)  + 
  theme(axis.text.x  = element_text(angle=60, vjust=0.5)) + 
  xlab("Date of MR Scan") +
  ylab("Timepoints Above Threshold (%)") + 
  ggtitle("Percent of TRs Above the 0.5mm FD Threshold Over Time") + 
  scale_x_date(date_labels = ("%b-%y"))
#maybe i ca nset the axes

p 
```

Once again, the horizontal red line indicates a cut-off of 50%. Points above the red line represent fMRI runs that will likely be unusable.

```{r}
#combo %>% ggplot(data=., mapping = aes(x=acq_site, y=fd_perc, fill=acq_id)) + geom_boxplot()+ facet_wrap(~acq_site) + ggtitle("Percent of Timepoints Above the 0.5mm Threshold") + xlab("Date of MR Scan") +ylab("Timepoints Above Threshold (%)") + labs(color="Scanner ID") + theme(axis.text.x  = element_text(angle=90, vjust=0.5))


combo %>% ggplot(data=., mapping = aes(x=acq_site, y=fd_perc, fill=acq_id)) + geom_boxplot() + geom_point(position=position_jitterdodge(jitter.width=0.1), alpha=0.2) + 
  ggtitle("Percent of Timepoints Above the 0.5mm Motion Threshold") + xlab("Site") +ylab("Timepoints Above Threshold (%)") + labs(fill="Scanner ID") + expand_limits(y=0) + expand_limits(y=100)  + geom_hline(yintercept = 50, linetype=1, colour="red") 


```


This is similar information as to above in a table format: 

```{r, results='asis'}
summary <- combo %>% group_by(acq_id) %>% summarize_at(vars("fd_mean", "fd_perc"),  funs(mean = median(., na.rm = TRUE), sd=sd(.))) %>% mutate(., fd_mean=paste0(round(fd_mean_mean, digits=2), '(', round(fd_mean_sd, digits=2),')'),fd_perc=paste0(round(fd_perc_mean,2), '(', round(fd_perc_sd,2),')')) %>% select(.,acq_id, fd_mean, fd_perc)

summary %>%
  kable(col.names=c("Scanner", "Median FD (SD)", "% FDs Above Threshold (SD)")) %>%
  kable_styling(bootstrap_options = c("striped"), full_width = F) %>% 
  column_spec(1, bold = T) %>% 
  group_rows("CU", 1, 2) %>%
 group_rows("LA", 3,3) %>% 
  group_rows("UP", 4, 5) %>% 
  group_rows("UT", 6, 7) %>%
  group_rows("WU", 8,9)
```


This demonstrates how the mean framewise displacement has changed over time: 
```{r}
xstart=as.Date("2017-11-18")
xend = as.Date("2018-12-14")

xstart2=xend
#xend2=as.Date(max(combo$mr_date))
xend2=as.Date("2019-02-20")

p <-  ggplot(data=combo, mapping = (aes(mr_date,fd_mean, col=acq_id)))+  
  annotate("rect", xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.5)  +
  annotate("rect", xmin = xstart2, xmax = xend2, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.8)  + 
  geom_point() + 
  geom_vline(xintercept=as.Date('2018-12-14'), linetype=3) + 
  facet_wrap(~scanner)  + 
  theme(axis.text.x  = element_text(angle=60, vjust=0.5)) + 
  xlab("Date of MR Scan") +
  ylab("Mean FD (mm)") + 
  ggtitle("Mean Framewise Displacement Over Time") + 
  scale_x_date(date_labels = ("%b-%y"))

p 
```



